<!--
    %\VignetteEngine{knitr::knitr}
    %\VignetteIndexEntry{ternaryplot graphical parameters: customise ternary plots}
-->

ternaryplot graphical parameters: customise ternary plots
============================================================

```{r echo=FALSE}
pv  <- sessionInfo( package = "ternaryplot" ) 
pv  <- pv[[ "otherPkgs" ]][["ternaryplot"]][["Version"]] 
now <- format( Sys.time(), "%Y-%m-%d %H:%M" ) 
```

> **Document information:**    <a id="table-document-information"></a>

|                  |                      | 
|:-----------------|---------------------:| 
| Author:          | **[Julien Moeys][]** | 
| Last build:      | `r now`              | 
| Package version: | `r pv`               | 

```{r echo=FALSE}
rm( pv, now )
```



Table of content    <a id="toc"></a>
------------------------------------------------------------

*   [Introduction](#section-intro). 
*   [Default graphical parameters](#section-default). 
*   [Pre-defined graphic styles with 'ternaryStyle'](#section-styles). 
*   [Graphical parameters for the device and plot regions](#section-region). 
*   [Graphical parameters for data points, text and lines](#section-data). 
*   [Graphical parameters for the axes](#section-axes). 
*   [Graphical parameters for the plot region's box and grid](#section-box-grid). 
*   [Graphical parameters for class polygons and labels](#section-classes). 
*   [Set the title and sub-title](#section-title). 
*   [Session information](#session-info). 



Introduction    <a id="section-intro"></a>
------------------------------------------------------------

Ternary plots produced with the package `ternaryplot` use as 
much as possible graphical parameters from R `graphics` 
library. These parameters can be retrieved and modified using 
the function `graphics::par`. Type `?par` in R command prompt 
to view the definition of these parameters (including many not 
relevant for `ternaryplot`).

For several graphical parameters in the package `ternaryplot` 
it was not possible to use `par`-parameters (because they don't 
exists or because it is not possible to use them unambiguously).
These parameters can be retrieved and modified using the 
function `ternaryplot::tpPar`. Type `?ternaryplot::tpPar` in 
R command prompt to view the definition of these parameters 
(including many non-graphic parameters).



Default graphical parameters    <a id="section-default"></a>
------------------------------------------------------------

Before showing how to change the default graphical parameters 
of ternary plots, here is an example showing a default 
ternary plot:

```{r}
library( 'ternaryplot' ) 

#   Backup the default 'graphics' parameters
.par <- par( no.readonly = TRUE )

#   Default ternary plot 
ternaryPlot( 
    s    = 'hypres', 
    main = 'Default graphical parameters', 
    sub  = 'package: ternaryplot' ) 
```



Pre-defined graphic styles with 'ternaryStyle'    <a id="section-styles"></a>
------------------------------------------------------------

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Choose a style
ternaryStyle( style = 'publication', margin = TRUE ) 

#   Default ternary plot 
ternaryPlot( 
    s    = 'hypres', 
    main = "Pre-defined graphic styles with 'ternaryStyle'" ) 
```



Graphical parameters for the device and plot regions    <a id="section-region"></a>
------------------------------------------------------------

> **Graphical parameters for the device and the plot region**:    <a id="table-region"></a>

| **.**           | **Device region** | **Plot region**    |
|:----------------|------------------:|-------------------:| 
| Draw:           | _n.r._            | _n.r._             | 
| Outline colour: | _n.r._            | _n.r._             | 
| Fill colour     | `par('bg')`       | `tpPar('plot.bg')` | 
| Expansion       | _n.r._            | _n.r._             | 
| Font            | _n.r._            | _n.r._             | 
| Font family     | _n.r._            | _n.r._             | 
| Line thickness  | _n.r._            | _n.r._             | 

> _n.r._: not relevant. The plot area is defined here as the 
> area inside the 3 axes. See `?graphics::par` and 
> `?ternaryplot::tpPar`. 



The example below shows how to modify these parameters:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Change (some) graphical parameters
par( bg = 'ivory' ) 
tpPar( plot.bg = 'peachpuff' ) 

#   Visualise the ternary plot
ternaryPlot( 
    s    = 'hypres', 
    main = 'Custom background colours', 
    sub  = 'package: ternaryplot' ) 
```

Notice that `ternaryplot` graphics internally call 
`ternaryWindow` to prepare the base graphic area, and that 
this function sets `par(pty='s')` (i.e. for a square 
plot-region).



Graphical parameters for data points, text and lines    <a id="section-data"></a>
------------------------------------------------------------

> **Graphical parameters for ternary data (points, text or lines)**:    <a id="table-data"></a>

| **.**           | **Points** (data) | **Text** (data) | **Lines** (data)  | 
|:----------------|------------------:|----------------:| -----------------:| 
| Draw (Y/N):     | (1)               | (2)             | (3)               | 
| Outline colour: | `col` (4)         | `col` (5)       | `col` (6)         | 
| Fill colour     | `bg`  (4)         | _n.r._          | _n.r._            | 
| Expansion       | `cex` (4)         | `cex` (5)       | _n.r._            |  
| Font            | _n.r._            | `font` (5)      | _n.r._            | 
| Font family     | _n.r._            | `family` (5)    | _n.r._            | 
| Line thickness  | `lwd` (4)         | _n.r._          | `lwd` (6)         |  

> (1) See argument `type` in `ternaryPlot` (`'n'` for nothing, 
> `'p'` for points). (2) Use `ternaryText` after `ternaryPlot`. 
> (3) Use `ternaryLines` after `ternaryPlot`. (4) Use arguments 
> `col`, `bg`, `cex` and `lwd` in `ternaryPlot` or 
> `ternaryPoints`. Use arguments `col`, `cex`, `font` and 
> `family` in `ternaryText`. (6) Use arguments `col`, 
> `lwd` (and also `lty`) in `ternarySegments`.
> _n.r._: not relevant. 



The examples below shows how to modify these parameters:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Load example dataset
data( 'textureDataset' )

#   Change (some) graphical parameters
#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'Custom graphical parameters for data points and text', 
    sub  = 'package: ternaryplot', 
    pch  = 21,           # (*) 
    col  = 'orangered4', # (*) 
    lwd  = 2,            # (*) 
    cex  = 1.5,          # (*) 
    bg   = 'orangered' ) # (*) 
# (*)   Passed to ternaryPoints(), then to points(), 
#       via "..."

ternaryText( 
    s      = s, 
    x      = textureDataset[ 1:20, ], 
    labels = as.character( 1:20 ), 
    col    = 'purple4', 
    pos    = 4,         # to the right 
    offset = 0.5, 
    font   = 2,         # bold
    family = 'serif' )  # (**)   
# (**)  Passed to text(), then to par(), via "..."
```

`ternaryLines` is not implemented yet, but `ternarySegments` 
can be used instead:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    main = 'Custom graphical parameters for data lines', 
    sub  = 'package: ternaryplot' ) 

ternarySegments( 
    s      = s, 
    from   = textureDataset[ 1:19, ], 
    to     = textureDataset[ 2:20, ], 
    col    = 'orangered4', 
    lwd    = 2, 
    lty    = 5 )   
```



Graphical parameters for the axes    <a id="section-axes"></a>
------------------------------------------------------------

> **Graphical parameters for ternary plot axes - part 1**:    <a id="table-axes-part1"></a>

| **.**           | **Axes lines**           | **Axes ticks**            | **Axes ticks labels**  | 
|:----------------|-------------------------:|--------------------------:| ----------------------:| 
| Draw (Y/N):     | (1)                      | (1)                       | (1)                    | 
| Outline colour: | `tpPar('axis.line.col')` | `tpPar('ticks.line.col')` | `par('col.axis')`      | 
| Fill colour     | _n.r._                   | _n.r._                    | _n.r._                 | 
| Expansion       | _n.r._                   | _n.r._                    | `par('cex.axis')`      |  
| Font            | _n.r._                   | _n.r._                    | `par('font.axis')`     | 
| Font family     | _n.r._                   | _n.r._                    | `par('family')`        | 
| Line thickness  | `tpPar('axis.line.lwd')` | `tpPar('ticks.line.lwd')` | _n.r._                 |  

> (1) See argument `axis` in `ternaryPlot`. See also `ternaryAxis`.
> _n.r._: not relevant. If `tpPar('axis.line.col')` or 
> `tpPar('ticks.line.col')` are `NULL`, `par('fg')` is used 
> instead. If `tpPar('axis.line.lwd')` or `tpPar('ticks.line.lwd')` 
> are `NULL`, `par('fg')` is used instead.



First an example showing how to draw a ternary triangle without 
axes (but with a box-frame around the plot region)

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    axes = FALSE, # NOTE: not 'axis'
    main = 'Ternary plot without axes', 
    sub  = 'package: ternaryplot' ) 
```



Then examples showing how to modify the axes graphical 
parameters:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Change (some) graphical parameters
par( col.axis = 'orangered4', cex.axis = 0.5, 
    font.axis = 2, family = 'serif' ) 

tpPar( axis.line.col = 'purple4', ticks.line.col = 'orangered4', 
    axis.line.lwd = 2, ticks.line.lwd = 3 )

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'Custom axes graphical parameters (1/2)', 
    sub  = 'package: ternaryplot' ) 
```



> **Graphical parameters for ternary plot axes - part 2**:    <a id="table-axes-part2"></a>

| **.**           | **Axes labels**          | **Axes arrow lines**      | 
|:----------------|-------------------------:|--------------------------:| 
| Draw (Y/N):     | (1)                      | `tpPar('arrows')` (2)     | 
| Outline colour: | `par('col.lab')`         | `tpPar('axis.line.col')`  | 
| Fill colour     | _n.r._                   | _n.r._                    | 
| Expansion       | `par('cex.lab')`         | _n.r._                    | 
| Font            | `par('font.lab')`        | _n.r._                    | 
| Font family     | `par('family')`          | _n.r._                    | 
| Line thickness  | _n.r._                   | `tpPar('axis.line.lwd')`  | 

> (1) See argument `axis` in `ternaryPlot`. (2) Set 
> `toPar('arrowsBreak'=FALSE)` to suppress the direction change 
> in the axes arrows (i.e. only parallel to the axes). 
> See also `ternaryAxis`. _n.r._: not relevant.  If 
> `tpPar('axis.line.col')` is `NULL`, `par('fg')` is used 
> instead. If `tpPar('axis.line.lwd')` is `NULL`, `par('lwd')` 
> is used instead. 



The example below show how to modify these parameters:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

#   Change (some) graphical parameters
par( col.lab = 'orangered4', cex.lab = 0.5, 
    font.lab = 2, family = 'serif' ) 

tpPar( axis.line.col = 'purple4', axis.line.lwd = 2, 
    arrowsBreak = FALSE )

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'Custom axes graphical parameters (2/2)', 
    sub  = 'package: ternaryplot' ) 
```



Graphical parameters for the plot region's box and grid    <a id="section-box-grid"></a>
------------------------------------------------------------

> **Graphical parameters for ternary plot box and grid**:    <a id="table-box-grid"></a>

| **.**           | **Box (over axes)**      | **Grid**                  | 
|:----------------|-------------------------:|--------------------------:| 
| Draw (Y/N):     | (1)                      | (2)                       | 
| Outline colour: | `tpPar('axis.line.col')` | `tpPar('grid.line.col')`  | 
| Fill colour     | `tpPar('plot.bg')`       | _n.r._                    | 
| Expansion       | _n.r._                   | _n.r._                    | 
| Font            | _n.r._                   | _n.r._                    | 
| Font family     | _n.r._                   | _n.r._                    | 
| Line thickness  | `tpPar('axis.line.lwd')` | `tpPar('grid.line.lwd')`  | 

> (1) See argument `frame.plot` in `ternaryPlot`. (2) The grid 
> cannot be suppressed at the moment, but consider setting 
> `tpPar('grid.line.col'=NA)` to make it invisible. If 
> `tpPar('grid.line.lwd')` is `NULL`, `par('lwd')` is used 
> instead. _n.r._: not relevant. 



First an example showing how the box around the plot and 
the grid can be suppressed from a ternary plot.

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

tpPar( grid.line.col = NA, axis.line.lwd = 2, 
    ticks.line.lwd = 2 )

#   Visualise the ternary plot
s <- ternaryPlot( 
    s          = 'default', 
    x          = textureDataset[ 1:20, ], 
    frame.plot = FALSE, 
    main       = 'Ternary plot without box-frame and grid', 
    sub        = 'package: ternaryplot' ) 
```

Notice that the absence of axes' title (axes' labels) is 
because the default ternary system is 'neutral' and does 
not have any pre-defined variables for each axis.



The example below show how to modify the graphical parameters 
for the plot-box and the grid:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

tpPar( axis.line.col = 'orangered4', axis.line.lwd = 2, 
    grid.line.col = 'gold', grid.line.lwd = 2, 
    plot.bg = 'peachpuff' ) 

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'Custom box and grid graphical parameters', 
    sub  = 'package: ternaryplot' ) 
```



Graphical parameters for class polygons and labels    <a id="section-classes"></a>
------------------------------------------------------------

> **Graphical parameters for ternary classification (polygons)**:    <a id="table-classes"></a>

| **.**           | **Class polygons**          | **Class labels**           | 
|:----------------|----------------------------:|---------------------------:| 
| Draw (Y/N):     | (1)                         | (1)                        | 
| Outline colour: | `tpPar('class.border.col')` | `tpPar('class.label.col')` | 
| Fill colour     | `tpPar('class.bg')`         | _n.r._                     | 
| Expansion       | _n.r._                      | `par('cex')`               | 
| Font            | _n.r._                      | `par('font')`              | 
| Font family     | _n.r._                      | `par('family')`            | 
| Line thickness  | `tpPar('class.border.lwd')` | _n.r._                     | 

> (1) Cannot be suppressed at the moment, but consider using 
> an emptty ternary classification instead. If 
> `tpPar('class.border.lwd')` is `NULL`, `par('lwd')` is used 
> instead. _n.r._: not relevant. 



The example below show how to modify these parameters:

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

tpPar( class.border.col = 'orangered4', 
    class.border.lwd = 2, class.label.col = 'orangered2', 
    class.bg = terrain.colors( n = 5 ) ) 

    #   Note: there must be as many colors in 'class.bg'
    #         as classes in the ternary classification 
    #         displayed! Extra colors are ignored.

par( font = 2, family = 'serif' ) 

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'Ternary-classes with custom graphical parameters', 
    sub  = 'package: ternaryplot' ) 
```



Set the title and sub-title    <a id="section-title"></a>
------------------------------------------------------------

```{r}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )

par( col.main = 'orangered4', col.sub = 'orangered2', 
    font.main = 1, font.sub = 3, cex.main = 1, 
    cex.sub = 1.5, family = 'serif' ) 

#   Visualise the ternary plot
s <- ternaryPlot( 
    s    = 'hypres', 
    x    = textureDataset[ 1:20, ], 
    main = 'A custom title', 
    sub  = 'package: ternaryplot' ) 
```


```{r echo=FALSE}
#   Reset graphical parameters
par( .par )
tpPar( reset = TRUE )
```



Session information    <a id="session-info"></a>
------------------------------------------------------------

```{r}
sessionInfo() 
```

<!-- List of links  -->
[Julien Moeys]:     http://julienmoeys.info "Julien Moeys homepage"

